maxent model avec xvalidation (year-based test training)

### 2019-04-30 src/maxent_avec_xvalidation.Rmd
binary complete - just now playing with a few bits

```{r pre-install & load packages, include=FALSE}
options(java.parameters = "-Xmx1g" )#needs to run before dismo loaded - boosts memory for java to 1g
packages <- c("dismo", "stringr", "rJava", "raster", "ecospat")
source("../src/ipak.R")
ipak(packages)
```

@@

@@

    #calculate minimal predicted area using ecospat
    aicetc$mpa_0.9_cum <- ecospat.mpa(samplepred$Cumulative.prediction, perc = 0.9) #90%

@@

    unlayer_yr <- subset(unlayer, ascyears == yr)
    unlayer_yr <- unlayer_yr[-c(1), ]
    unlayer_yr[] <- lapply(unlayer_yr, as.character) #from https://stackoverflow.com/questions/2851015/convert-data-frame-columns-from-factors-to-characters
    stackdeplst <- list()
    stackdep <- subset(unlayer_yr, select = c(stackname, ascdepths))
    stackdeplst[[year]] <- stackdep
    #write.csv(stackdep, paste0("../output/maxent/predictions/stackdepthlist_", yr, ".csv"), row.names = FALSE)
    aea <- raster("../output/env/aea.tif")

   for (td in 1:nrow(unlayer_yr)){
      temp_depth <- raster(unlayer_yr$temp_depth[td])
      crs(temp_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(temp_depth) <- "temp_depth"
      temp_depth <- projectRaster(temp_depth, aea)
      chl_surface <- raster(unlayer_yr$chl_surface[td])
      crs(chl_surface) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(chl_surface) <- "chl_surface"
      chl_surface <- projectRaster(chl_surface, aea)
      o2_depth <- raster(unlayer_yr$o2_depth[td])
      crs(o2_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(o2_depth) <- "o2_depth"
      o2_depth <- projectRaster(o2_depth, aea)
      salinity_depth <- raster(unlayer_yr$salinity_depth[td])
      crs(salinity_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(salinity_depth) <- "salinity_depth"
      salinity_depth <- projectRaster(salinity_depth, aea)
      nao_sample <- raster(unlayer_yr$nao_sample[td])
      crs(nao_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(nao_sample) <- "nao_sample"
      nao_sample <- projectRaster(nao_sample, aea)
      nao_winter <- raster(unlayer_yr$nao_winter[td])
      crs(nao_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
      names(nao_winter) <- "nao_winter"
      nao_winter <- projectRaster(nao_winter, aea)
      stackn <- unlayer_yr$stackname[td]
      stack <- stack(temp_depth, salinity_depth, o2_depth, chl_surface, nao_sample, nao_prev, nao_winter, amo_sample, amo_winter)
      #stackSave(test,unlayer_yr$stackname[td])
      pred <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mthna, "/", "cloglog_", stackn, ".asc"), overwrite=TRUE, args='outputformat=cloglog') 
      predcum <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mthna, "/", "cumu_", stackn, ".asc"), overwrite=TRUE, args=c('outputformat=Cumulative')) 
      prplot <- plot(pred, png(paste0("../output/maxent/predictions/", mthna, "/", "cloglog_", stackn, ".png")))
      dev.off() # stops automatic saving of the plot to a png

@@

@@

```{r avg asc layers cloglog}
asc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "cloglog*.asc", full.names = TRUE)
depthlayers <- data.frame(cbind(unique(stdep$ascdepths)))
names(depthlayers)[names(depthlayers) == "cbind.unique.stdep.ascdepths.."] <- "depth"
depthlayers$depth <- as.character(depthlayers$depth)

@@

for (x in 1:nrow(depthlayers)){
  asclay <- depthlayers[x, ]
  asclay <- asclay[!is.na(asclay)]
  asclay <- as.data.frame(rbind(asclay))
  asclay$V1 <- as.character(asclay$V1) #v1 corresponds to depth
  asclay$V2 <- as.character(asclay$V2) #v2 to end corresponds to year
  asclay$V3 <- as.character(asclay$V3)
  asclay$V4 <- as.character(asclay$V4)
  asclay$V5 <- as.character(asclay$V5)
  asclay$V6 <- as.character(asclay$V6)
  asclay$V7 <- as.character(asclay$V7)
  asclay$V8 <- as.character(asclay$V8)
  asclay$V9 <- as.character(asclay$V9)
  asclay$V10 <- as.character(asclay$V10)
  asclay$V11 <- as.character(asclay$V11)
  asclay$V12 <- as.character(asclay$V12)
  asclay$V13 <- as.character(asclay$V13)
  asclay$V14 <- as.character(asclay$V14)
  asclay$V15 <- as.character(asclay$V15)

  asclaydep <- asclay$V1

  asc1998 <- raster(asclay$V2)
  asc1999 <- raster(asclay$V3)
  asc2000 <- raster(asclay$V4)
  asc2001 <- raster(asclay$V5)
  asc2002 <- raster(asclay$V6)
  asc2003 <- raster(asclay$V7)
  asc2004 <- raster(asclay$V8)
  asc2005 <- raster(asclay$V9)
  asc2006 <- raster(asclay$V10)
  asc2007 <- raster(asclay$V11)
  asc2008 <- raster(asclay$V12)
  asc2009 <- raster(asclay$V13)
  asc2010 <- raster(asclay$V14)
  asc2011 <- raster(asclay$V15)

  stkasc <- stack(asc1998, asc1999, asc2000, asc2001, asc2002, asc2003, asc2004, asc2005, asc2006, asc2007, asc2008, asc2009, asc2010, asc2011)
  avg <- calc(stkasc, fun = mean)

  writeRaster(avg, (paste0("../output/maxent/predictions/", mthna, "/avg_", asclaydep, ".asc")), overwrite=TRUE, args='outputformat=cloglog')
  avgplot <- plot(avg, png(paste0("../output/maxent/predictions/", mthna, "/avg_", asclaydep, ".png"))) 
  dev.off()
}

ascdepth_list <- list.files(paste0("../output/maxent/predictions/", mthna), glob2rx(pattern = "avg_*.asc"), full.names = TRUE)

clog_asc120  <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_120.asc"))
clog_asc41.18 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_41.18.asc"))
clog_asc1.55586 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1.55586.asc"))
clog_asc1045.85 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1045.85.asc"))
clog_asc108.03 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_108.03.asc"))
clog_asc11.7737 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_11.7737.asc"))
clog_asc1151.99 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1151.99.asc"))
clog_asc1265.86 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1265.86.asc"))
clog_asc13.991 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_13.991.asc"))
clog_asc133.076 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_133.076.asc"))
clog_asc1387.38 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1387.38.asc"))
clog_asc147.406 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_147.406.asc"))
clog_asc16.5253 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_16.5253.asc"))
clog_asc163.165 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_163.165.asc"))
clog_asc180.55 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_180.55.asc"))
clog_asc19.4298 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_19.4298.asc"))
clog_asc199.79 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_199.79.asc"))
clog_asc2.66768 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_2.66768.asc"))
clog_asc22.7576 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_22.7576.asc"))
clog_asc221.141 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_221.141.asc"))
clog_asc244.891 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_244.891.asc"))
clog_asc26.5583 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_26.5583.asc"))
clog_asc271.356 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_271.356.asc"))
clog_asc3.85628 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_3.85628.asc"))
clog_asc30.8746 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_30.8746.asc"))
clog_asc300.888 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_300.888.asc"))
clog_asc333.863 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_333.863.asc"))
clog_asc35.7402 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_35.7402.asc"))
clog_asc370.689 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_370.689.asc"))
clog_asc411.794 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_411.794.asc"))
clog_asc457.626 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_457.626.asc"))
clog_asc47.2119 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_47.2119.asc"))
clog_asc5.14036 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_5.14036.asc"))
clog_asc508.64 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_508.64.asc"))
clog_asc53.8506 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_53.8506.asc"))
clog_asc565.292 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_565.292.asc"))
clog_asc6.54303 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_6.54303.asc"))
clog_asc61.1128 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_61.1128.asc"))
clog_asc628.026 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_628.026.asc"))
clog_asc69.0217 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_69.0217.asc"))
clog_asc697.259 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_697.259.asc"))
clog_asc77.6112 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_77.6112.asc"))
clog_asc773.368 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_773.368.asc"))
clog_asc8.09252 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_8.09252.asc"))
clog_asc856.679 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_856.679.asc"))
clog_asc86.9294 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_86.9294.asc"))
clog_asc9.82275 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_9.82275.asc"))
clog_asc947.448 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_947.448.asc"))
clog_asc97.0413 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_97.0413.asc"))

stackavg <- stack(clog_asc120, clog_asc41.18, clog_asc1.55586, clog_asc1045.85, clog_asc108.03, clog_asc11.7737, clog_asc1151.99, clog_asc1265.86, clog_asc13.991, clog_asc133.076, clog_asc1387.38, clog_asc147.406, clog_asc16.5253, clog_asc163.165, clog_asc180.55, clog_asc19.4298, clog_asc199.79, clog_asc2.66768, clog_asc22.7576, clog_asc221.141, clog_asc244.891, clog_asc26.5583, clog_asc271.356, clog_asc3.85628, clog_asc30.8746, clog_asc300.888, clog_asc333.863, clog_asc35.7402, clog_asc370.689, clog_asc411.794, clog_asc457.626, clog_asc47.2119, clog_asc5.14036, clog_asc508.64, clog_asc53.8506, clog_asc565.292, clog_asc6.54303, clog_asc61.1128, clog_asc628.026, clog_asc69.0217, clog_asc697.259, clog_asc77.6112, clog_asc773.368, clog_asc8.09252, clog_asc856.679, clog_asc86.9294, clog_asc9.82275, clog_asc947.448, clog_asc97.0413)

stsum <- sum(stackavg, na.rm = TRUE)
  writeRaster(stsum, (paste0("../output/maxent/predictions/", mthna, "/clog_sum_", mthna, ".asc")), overwrite=TRUE)
  sumplot <- plot(stsum, png(paste0("../output/maxent/predictions/", mthna, "/clog_sum_", mthna, ".png"))) 
  dev.off()
stavg <- calc(stackavg, fun = mean, na.rm = TRUE)
  writeRaster(stavg, (paste0("../output/maxent/predictions/", mthna, "/clog_avg_", mthna, ".asc")), overwrite=TRUE)
  avgplot <- plot(stavg, png(paste0("../output/maxent/predictions/", mthna, "/clog_avg_", mthna, ".png"))) 
  dev.off()

@@

#binary maps

```{r binary map using maxsumsenspec}
cumasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^cumu_.*.asc", full.names = TRUE)

@@

>>>>

```{r pre-install & load packages, include=FALSE}
options(java.parameters = "-Xmx1g" )#needs to run before dismo loaded - boosts memory for java to 1g
packages <- c("dismo", "stringr", "rJava", "raster", "ecospat", "bossMaps")
source("../src/ipak.R")
ipak(packages)
```

@@

      aicetc$train_AUC <- trainauc #training only AUC - not tested
      aicetc$no_trainng_points <- notrainingpoints #numberof points used for training
      aicetc$test_coefficients <- nocoeff #this is to calculate the AIC
      aicetc$test_sum.of.logs <- sumsamplepred #this is to calculate the AIC
      aicetc$test_AUC <- evauc #testing AUC
      aicetc$test_evcor <- evcor #testing correlation coefficient
      aicetc$test_evpcor <- evpcor # testing p-value for correlation coefficient
      aicetc$test_nopresence <- evnp 
      aicetc$test_nobackground <- evna
      aicetc$test_maxsumsenspec <- evmaxtprtnr
      aicetc$test_maxkap <- evmaxkap
      aicetc$test_no_omission <- evno_omission
      aicetc$test_prevalence <- evprevalence
      aicetc$test_equal_sens_spec <- evequal_sens_spec
      aicetc$test_sensitivity <- evsensitivity
      aicetc$test_bic <- bic #not sure this is correct...
      aicetc$beta <- betaval

@@

    #calculate minimal predicted area using ecospat
    aicetc$mpa_0.9_cum <- ecospat.mpa(samplepred$Cumulative.prediction, perc = 0.9) #90%
    aicetc$mpa_0.9_clog <- ecospat.mpa(samplepred$Cloglog.prediction, perc = 0.9) #90%
    aicetc$year <- yr

@@

    unlayer_yr <- subset(unlayer, ascyears == yr)
    unlayer_yr <- unlayer_yr[-c(1), ]
    unlayer_yr[] <- lapply(unlayer_yr, as.character) #from https://stackoverflow.com/questions/2851015/convert-data-frame-columns-from-factors-to-characters
    stackdeplst <- list()
    stackdep <- subset(unlayer_yr, select = c(stackname, ascdepths))
    stackdeplst[[year]] <- stackdep
    write.csv(stackdep, paste0("../output/maxent/predictions/stackdepthlist_", yr, ".csv"), row.names = FALSE)
    aea <- raster("../output/env/aea.tif")

   for (td in 1:nrow(unlayer_yr)){
      temp_depth <- raster(unlayer_yr$temp_depth[td])
      crs(temp_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(temp_depth) <- "temp_depth"
      temp_depth <- projectRaster(temp_depth, aea)
      chl_surface <- raster(unlayer_yr$chl_surface[td])
      crs(chl_surface) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(chl_surface) <- "chl_surface"
      chl_surface <- projectRaster(chl_surface, aea)
      o2_depth <- raster(unlayer_yr$o2_depth[td])
      crs(o2_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(o2_depth) <- "o2_depth"
      o2_depth <- projectRaster(o2_depth, aea)
      salinity_depth <- raster(unlayer_yr$salinity_depth[td])
      crs(salinity_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(salinity_depth) <- "salinity_depth"
      salinity_depth <- projectRaster(salinity_depth, aea)
      nao_sample <- raster(unlayer_yr$nao_sample[td])
      crs(nao_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(nao_sample) <- "nao_sample"
      nao_sample <- projectRaster(nao_sample, aea)
      nao_winter <- raster(unlayer_yr$nao_winter[td])
      crs(nao_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
      names(nao_winter) <- "nao_winter"
      nao_winter <- projectRaster(nao_winter, aea)
      stackn <- unlayer_yr$stackname[td]
      stack <- stack(temp_depth, salinity_depth, o2_depth, chl_surface, nao_sample, nao_winter)
      #stackSave(test,unlayer_yr$stackname[td])
      pred <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mthna, "/", "cloglog_", stackn, ".asc"), overwrite=TRUE, args='outputformat=cloglog') 
      predraw <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mthna, "/", "raw_", stackn, ".asc"), overwrite=TRUE, args=c('outputformat=raw')) 
      prplot <- plot(pred, png(paste0("../output/maxent/predictions/", mthna, "/", "cloglog_", stackn, ".png")))
      dev.off() # stops automatic saving of the plot to a png

@@

#binary maps - individual layers

for some reason the cumulative output comes empty... using bossMaps
```{r raw to cumulative}
rawasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^raw_.*.asc", full.names = TRUE) 
for (raw in 1:length(rawasc_list)){
  asc <- raster(rawasc_list[raw])
  ascname <- names(asc)
  ascname <- str_sub(ascname, start = 5)
  cumasc <- cumulative(asc, threshold = NULL)
  writeRaster(cumasc, paste0("../output/maxent/predictions/", mthna, "/", "cum_", ascname, ".asc"), overwrite = TRUE)
}
```

```{r set up threshold binary function}
binmpa <- function(x) {
  ifelse(x <=  mpa_thr, 0,
  ifelse(x >  mpa_thr, 1, NA))
  }
```
cum_mpathr <- calc(cumasc, fun = rc)
plot(cum_mpathr)
tenper_thr <- 20.338725

```{r binary maps on individual layers}
cumasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^cum_.*.asc", full.names = TRUE)

for (modtime in 1:nrow(modelmonth)){
  timeslice <- modelmonth[modtime]
  print(timeslice)
  yr <- str_sub(timeslice, -4)
  print(yr)
  thresholds <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
  mpa_thr <- thresholds$mpa_0.9_cum
  print(mpa_thr)
  for (cum in 1:length(cumasc_list)){
    cumascyr <- str_sub(cumasc_list[cum], start = 40, end = 43) #selects the year from the file name
    if(yr == cumascyr){
      cumasc <- raster(cumasc_list[cum])
      ascname <- names(cumasc)
      print(ascname)
      cum_mpathr <- calc(cumasc, fun = binmpa)
      mpa_plot <- plot(cum_mpathr, png(paste0("../output/maxent/predictions/", mthna, "/", "binary_mpa_", ascname, ".png")))
      dev.off() # stops automatic saving of the plot to a png
      writeRaster(cum_mpathr, paste0("../output/maxent/predictions/", mthna, "/", "binary_mpa_", ascname, ".asc"), overwrite = TRUE)
    }
  }
}

@@

```{r avg asc layers cloglog}
asc_list <- list.files(paste0("../output/maxent/predictions/", mthna, "/"), pattern = "^cloglog.*asc", full.names = TRUE)
depthlayers <- data.frame(cbind(unique(stdep$ascdepths)))
names(depthlayers)[names(depthlayers) == "cbind.unique.stdep.ascdepths.."] <- "depth"
depthlayers$depth <- as.character(depthlayers$depth)

@@

for (x in 1:nrow(depthlayers)){
  asclay <- depthlayers[x, ]
  asclay <- asclay[!is.na(asclay)]
  asclay <- as.data.frame(rbind(asclay))
  asclay$V1 <- as.character(asclay$V1)
  asclay$V2 <- as.character(asclay$V2)
  asclay$V3 <- as.character(asclay$V3)
  asclay$V4 <- as.character(asclay$V4)
  asclay$V5 <- as.character(asclay$V5)
  asclaydep <- asclay$V1

  asc2003 <- raster(asclay$V2)
  asc2004 <- raster(asclay$V3)
  asc2005 <- raster(asclay$V4)
  asc2006 <- raster(asclay$V5)

  stkasc <- stack(asc2003, asc2004, asc2005, asc2006)
  avg <- calc(stkasc, fun = mean)

  writeRaster(avg, paste0("../output/maxent/predictions/", mthna, "/avg_cloglog_", asclaydep, ".asc"), overwrite=TRUE)
  avgplot <- plot(avg, png(paste0("../output/maxent/predictions/", mthna, "/avg_cloglog_", asclaydep, ".png"))) 
  dev.off()
}
```

```{r avg asc layers cumulative}
asc_list <- list.files(paste0("../output/maxent/predictions/", mthna, "/"), pattern = "^cum.*asc", full.names = TRUE)
depthlayers <- data.frame(cbind(unique(stdep$ascdepths)))
names(depthlayers)[names(depthlayers) == "cbind.unique.stdep.ascdepths.."] <- "depth"
depthlayers$depth <- as.character(depthlayers$depth)

for (d in 1:nrow(depthlayers)){
  dep <- depthlayers$depth[d]
  for (a in 1:length(asc_list)){
    asc <- asc_list[a]
    if (grepl(dep, asc)){
      depthlayers[d , paste0("a", colno)] <- asc
        colno <- colno+1
    }
  }
}
write.csv(depthlayers, (paste0("../output/maxent/predictions/", mthna, "/cum_asclayerbydepth.csv")), row.names = TRUE)

for (x in 1:nrow(depthlayers)){
  asclay <- depthlayers[x, ]
  asclay <- asclay[!is.na(asclay)]
  asclay <- as.data.frame(rbind(asclay))
  asclay$V1 <- as.character(asclay$V1)
  asclay$V2 <- as.character(asclay$V2)
  asclay$V3 <- as.character(asclay$V3)
  asclay$V4 <- as.character(asclay$V4)
  asclay$V5 <- as.character(asclay$V5)
  asclaydep <- asclay$V1

  asc2003 <- raster(asclay$V2)
  asc2004 <- raster(asclay$V3)
  asc2005 <- raster(asclay$V4)
  asc2006 <- raster(asclay$V5)

  stkasc <- stack(asc2003, asc2004, asc2005, asc2006)
  avg <- calc(stkasc, fun = mean)

  writeRaster(avg, paste0("../output/maxent/predictions/", mthna, "/avg_cum_", asclaydep, ".asc"), overwrite=TRUE)
  avgplot <- plot(avg, png(paste0("../output/maxent/predictions/", mthna, "/avg_cum_", asclaydep, ".png"))) 
  dev.off()
}

@@

```{r binary maps on averaged rasters}
avgcumasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^avg_cum.*.asc", full.names = TRUE)
thresholds <- read.csv(paste0("../output/maxent/models/", mthna, "avgAIC.csv"), header = TRUE)
thresholds <- tail(thresholds, 1) #subsset the last row
mpa_thr <- thresholds$mpa_0.9_cum
#print(mpa_thr)

for (avgcum in 1:length(avgcumasc_list)){
    cumasc <- raster(avgcumasc_list[avgcum])
    ascname <- names(cumasc)
    #print(ascname)
    cum_mpathr <- calc(cumasc, fun = binmpa)
    mpa_plot <- plot(cum_mpathr, png(paste0("../output/maxent/predictions/", mthna, "/", "binary_mpa_", ascname, ".png")))
    dev.off() # stops automatic saving of the plot to a png
    writeRaster(cum_mpathr, paste0("../output/maxent/predictions/", mthna, "/", "binary_mpa_", ascname, ".asc"), overwrite = TRUE, datatype="LOG1S")
}

@@

now to stack average binary maps at depth into one 2D representation...

```{r}

#this below is to get an average score for all layers per year. probably not what you want
ascdepth_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_mpa_avg_cum_.*asc", full.names = TRUE)

binaryavgstack <- stack(ascdepth_list)

stsum <- sum(binaryavgstack, na.rm = TRUE)
  writeRaster(stsum, (paste0("../output/maxent/predictions/", mthna, "/binary_sum_", mthna, ".asc")), overwrite=TRUE)
  sumplot <- plot(stsum, png(paste0("../output/maxent/predictions/", mthna, "/binary_sum_", mthna, ".png"))) 
  dev.off()
stavg <- calc(binaryavgstack, fun = mean, na.rm = TRUE)
  writeRaster(stavg, (paste0("../output/maxent/predictions/", mthna, "/binary_avg_", mthna, ".asc")), overwrite=TRUE)
  avgplot <- plot(stavg, png(paste0("../output/maxent/predictions/", mthna, "/binary_avg_", mthna, ".png"))) 
  dev.off()
